Gravitational Lens Recovery with \Glass: How to measure the mass profile and shape of a lens

### 2013-06-12 limits/ms.tex
further improvements on the intro along the lines suggested by P

{\bf JR. Lots of references missing here. Need to be added. But this should get us started.}\\
{\bf JR. I updated the abstract to lower the emphasis on non-parametric modelling and the connection with PixeLens. I will perform similar edits on the introduction too (c.f. comment from Prasenjit.}

Since the discovery of the first gravitational lens \citep{1979Natur.279..381W}, lensing has become an increasingly important probe of the distribution of mass in the Universe \citep{1937ApJ....86..217Z,2012arXiv1206.1225A}. Strong lensing in particular provides a powerful probe of the mass distribution at the heart of galaxies and galaxy clusters given good enough data \citep{2010CQGra..27w3001B,2007ApJ...667..645R,2006ApJ...652L...5S,2009ApJ...690..154S}. This in turn provides valuable constraints on the nature of dark matter \citep{2006ApJ...652L...5S,2013ApJ...765...25N}, on cosmological models \citep{2010Sci...329..924J, 2008ApJ...679...17C}, and on galaxy formation models \citep{2012MNRAS.424..104L}. To date, some $\sim$400 strong lenses are known\footnote{{\tt http://admin.masterlens.org/}.}, but with the advent of large ground and space based surveys, this number is set to explode, with thousands of lenses expected to be discovered over the next ten years \citep{2012arXiv1206.1225A, 2010AAS...21540115M, 2004NewAR..48.1085K}.

The majority of lens modelling efforts to date have focussed on `parametric' modelling, where simple theoretically motived models are fit using maximum likelihood methods to the data \citep{2011A&ARv..19...47K, 1993A&A...273..367K, 2010GReGr..42.2151K}. This is useful in that a simple model provides valuable physical insight. However, as data quality improves and the questions we ask of these data become more refined, it is important to re-asses the impact of our model assumptions. `Non-parameteric' methods -- that assume many more parameters than the available data constraints -- are invaluable in this respect. In non-parametric mass modelling, the lens inversion problem is deliberately chosen to be under-constrained \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2010ApJ...723.1678C,2006MNRAS.367.1209L,2013arXiv1304.2393S}. Instead of finding one `best-fit' model, we must then search for many models consistent with the data, exploring degeneracies between different but plausible\footnotemark\ mass models \citep{2000AJ....120.1654S,2006ApJ...653..936S,2008MNRAS.386..307L}. The key advantages of non-parametric modelling are that we can: (i) explore model degeneracies and marginalise over them; and (ii) we can determine what type and quality of data (including data from stellar population synthesis modelling, stellar kinematics, or other indepedent probes of the mass) are required to measure parameters of interest.

\footnotetext{The word {\it plausible} encodes the key hidden gremlin in non-parametric methods: constraints extra to the available data (priors) are typically required on the solution space in order to make it finite. However, even rather restrictive priors are far less restrictive than the implicit constraints inherent in parametric mass modelling.}

In this paper, we present a new non-parametric strong lens modelling framework \Glass\ that builds on an earlier code \PixeLens\ \citep{1997MNRAS.292..148S,2008ApJ...679...17C}. Like \PixeLens, we use a grid of mass pixels to model the lens, searching through many models consistent with the data to marginalise over model uncertainties. However, \Glass\ improves on \PixeLens\ in a number of important respects:

>>>>

{\bf JR. Lots of references missing here. Need to be added. But this should get us started.}\\

Since the discovery of the first gravitational lens \citep{1979Natur.279..381W}, lensing has become an increasingly important probe of the distribution of mass in the Universe \citep{1937ApJ....86..217Z,2012arXiv1206.1225A}. Strong lensing in particular provides a powerful probe of the mass distribution at the heart of galaxies and galaxy clusters given good enough data \citep{2010CQGra..27w3001B,2007ApJ...667..645R,2006ApJ...652L...5S,2009ApJ...690..154S}. This in turn provides valuable constraints on the nature of dark matter \citep{2006ApJ...652L...5S,2013ApJ...765...25N}, on cosmological models \citep{2010Sci...329..924J, 2008ApJ...679...17C}, and on galaxy formation models \citep{2012MNRAS.424..104L}. To date, some $\sim$400 strong lenses are known\footnote{{\tt http://admin.masterlens.org/}.}, but with the advent of large ground and space based surveys, many thousands of lenses are expected to be discovered over the next ten years \citep{2012arXiv1206.1225A, 2010AAS...21540115M, 2004NewAR..48.1085K}.

With this coming explosion in lens data, it is timely to re-asses the impact of model degeneracies on quantities of interest like the circularly averaged density profile of a lens or its shape. Some significant work has already proceeded in this direction in the literature. XXX found that YYY. However, most of these studies employed simplified models that assume a specific analytic form for the lens and/or connect the shape of the underlying dark matter halo to its parent galaxy. Notable exceptions are XXX. As the quality and quantity of data improve, we can afford relax some of our historical model assumptions and attempt to actually {\it image} the distribution of mass in lensing galaxies and clusters. We refer to this as `non-parametric' mass modelling, where `non-' means that we deliberately employ far more model parameters than data constraints. Searching through our model parameter space, we can then explore model degeneracies and determine which data best break them.

%The majority of lens modelling efforts to date have focussed on `parametric' modelling, where simple theoretically motived models are fit using maximum likelihood methods to the data \citep{2011A&ARv..19...47K, 1993A&A...273..367K, 2010GReGr..42.2151K}. This is useful in that a simple model provides valuable physical insight. However, as data quality improves and the questions we ask of these data become more refined, it is important to re-asses the impact of our model assumptions. `Non-parameteric' methods -- that assume many more parameters than the available data constraints -- are invaluable in this respect. In non-parametric mass modelling, the lens inversion problem is deliberately chosen to be under-constrained \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2010ApJ...723.1678C,2006MNRAS.367.1209L,2013arXiv1304.2393S}. Instead of finding one `best-fit' model, we must then search for many models consistent with the data, exploring degeneracies between different but plausible\footnotemark\ mass models \citep{2000AJ....120.1654S,2006ApJ...653..936S,2008MNRAS.386..307L}. The key advantages of non-parametric modelling are that we can: (i) explore model degeneracies and marginalise over them; and (ii) we can determine what type and quality of data (including data from stellar population synthesis modelling, stellar kinematics, or other indepedent probes of the mass) are required to measure parameters of interest.
%\footnotetext{The word {\it plausible} encodes the key hidden gremlin in non-parametric methods: constraints extra to the available data (priors) are typically required on the solution space in order to make it finite. However, even rather restrictive priors are far less restrictive than the implicit constraints inherent in parametric mass modelling.}

%that builds on an earlier code \PixeLens\ \citep{1997MNRAS.292..148S,2008ApJ...679...17C}. Like \PixeLens, we

In this paper, we present a new non-parametric strong lens modelling framework \Glass\ that uses an adaptive grid of mass pixels to model the lens, searching through many models consistent with the data to marginalise over model uncertainties. \Glass\ shares some aspects with an earlier code \PixeLens\ \citep{1997MNRAS.292..148S,2008ApJ...679...17C}, but has important new functionality:
